Entropy-SGD: Biasing Gradient Descent Into Wide Valleys

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
wip

The authors in~\citet{hochreiter1997flat} introduce a similar technique as our work, they introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization. \fix{Their analysis of the generalization performance (see their Sec. A1) necessitates a few unrealistic assumptions whereas in our analysis, as will be shown, we use uniform stability~\citet{bousquet2002stability} to obtain a precise characterization of generalization.}{one sentence about the similarities} Furthermore, as they mention, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems. Their formulation depends on Hessian-vector products. It is unclear how this method could be scaled to modern deep networks. Motivated by the same final goal, viz., flat local minima, our approach eliminates these extraneous and difficult to control hyper-parameters and instead results in a scalable algorithm. Also recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper. The authors estimate the loss in a neighborhood of the weights to argue that small batch size in SGD (i.e., larger gradient noise) generalizes better than large mini-batches and also results in significantly flatter minima.

@@

\pc{If we focus on the inner loop of the algorithm, SGLD updates compute the average gradient (with Langevin noise) in a neighborhood of the parameters while maintaining the Polyak average of the new parameters. Such an interpretation is very close to averaged SGD of~\citet{polyak1992acceleration} and~\citet{bottou2012stochastic} and worth further study. Moreover, our experiments show that $\entropysgd$ trains significantly faster than SGD for recurrent networks with relatively minor gains in \fix{wall-clock time for CNNs}{}. Estimating the gradient of local entropy cheaply with few SGLD iterations, or more interestingly, by using a smaller network in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.}{don't want to give away the farm}

>>>>

Motivated by the same final goal, viz., flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} \fix{(their analysis necessitates a few unrealistic assumptions, see their Sec. A1)}{we might consider deleting this} that this leads to improved generalization. However, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems, \textcolor{red}{as discussed by the authors.} Their formulation finally depends on Hessian-vector products. Driven by the same goal of encouraging the optimizer to favor flat regions in the opimization landscape, we however formulate a different objective function based on local entropy and obtain a more scalable algorithm which does not rely on extraneous and difficult to control hyper-parameters. \fix{Despite fundamental differences, both approaches... }{one sentence about the similarities} Also recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper. The authors estimate the loss in a neighborhood of the weights to argue that small batch size in SGD (i.e., larger gradient noise) generalizes better than large mini-batches and also results in significantly flatter minima.

@@

If we focus on the inner loop of the algorithm, SGLD updates compute the average gradient (with Langevin noise) in a neighborhood of the parameters while maintaining the Polyak average of the new parameters. Such an interpretation is very close to averaged SGD of~\citet{polyak1992acceleration} and~\citet{bottou2012stochastic} and worth further study. Our experiments show that while $\entropysgd$ trains significantly faster than SGD for recurrent networks, it gets relatively minor gains in terms of wall-clock time for CNNs. Estimating the gradient of local entropy cheaply with few SGLD iterations, or by using a smaller network to estimate the gradient in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.
